---
title: "R4DS 12.2.1.2"
author: "ses"
date: "5/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setup

```{r message = FALSE}
library(tidyverse)
```


# Aufgabe

[Quelle](https://r4ds.had.co.nz/tidy-data.html#tidy-data-1)

>   Compute the rate for table2, and table4a + table4b. You will need to perform four operations:

a. Extract the number of TB cases per country per year.
b. Extract the matching population per country per year.
c. Divide cases by population, and multiply by 10000.
d. Store back in the appropriate place.


>    Which representation is easiest to work with? Which is hardest? Why?


## table2

```{r}
data(table2)
```


```{r}
head(table2)
```

### a. TB cases per country per year

```{r}
table2 %>% 
  filter(type == "cases") %>% 
  arrange(country, year)
```


### b.  Extract the matching population per country per year.

```{r}
table2 %>% 
  filter(type == "population") %>% 
  arrange(country, year)
```
### c./d. Divide cases by population, and multiply by 10000.

```{r}
table2a <- 
  table2 %>% 
  pivot_wider(names_from = type, values_from = count) %>% 
  mutate(rate = cases / population * 1000)

table2a
```


## table4a und table4b

```{r}
table4a %>% 
  head()
```


```{r}
table4b %>% 
  head()
```


```{r}
table4a %>% 
  full_join(table4b, by = "country") %>% 
  pivot_longer(cols = -country,
               names_to = "year",
               values_to = "values") %>% 
  separate(col = year, into = c("year", "variable")) %>% 
  mutate(variable = case_when(
    variable == "x" ~ "cases",
    variable == "y" ~ "population"
  ))  %>% 
  pivot_wider(names_from = variable,
              values_from = values) %>% 
  mutate(rate = cases/population * 1000)
```

