---
title: '2021-05-07'
author: "ses"
date: "5/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


# Vorbereitung

```{r message = FALSE}
library(tidyverse)
```


```{r echo = FALSE}
data("diamonds")
```

`data()` lädt Datensätze aber nur aus Paketen, die gestartet sind (d.h. mit `library()`)!

# Aufgaben

## 7.3.4.1

Explore the *distribution* of *each* of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

- Univariate Betrachtung: Jede Variable einzeln
- Verteilung (metr. Variable: Histogramm!)

```{r eval = FALSE}
ggplot(diamonds) +
  aes(x = carat) +
  geom_histogram() +
  xlim(3,9) +
  ylim(0, 10)
```


## Ausreißer: ungewöhnliche Werte



```{r fig.align='center', out.width="70%", fig.height=3}
ggplot(diamonds) +
  aes(x = carat) +
  geom_histogram() 
```


Wie kann man die Ausreißer besser sehen?

Boxplot! 

Erstellen Sie eine Variante mit Boxplot!


```{r}
ggplot(diamonds) +
  aes(x = carat) +
  geom_boxplot() 
```



# ifelse


```{r}
diamonds %>% 
  summarise(median(carat))

diamond2 <-
  diamonds %>% 
  mutate(grosser_diamand = ifelse(carat >= 0.7, 
                                  "gross", NA)) %>% 
  select(grosser_diamand, carat)
```


# filter und group_by

Mittlere Preis pro Schlliffart, aber nur für große Diamanten (>0.7 carat)!


```{r}
diamonds %>% 
  filter(carat > 0.7) %>% 
  group_by(cut) %>% 
  summarise(price_per_cut_mean = mean(price))
```



# Modulo-Rechnen


```{r}
2021 %/% 100 # ganzzahliger Teil des Ergebnisses

715 %/% 100

7/2  

7 %% 2  # Rest (Modulo)
```




# Fehlende Werte

```{r}
diamonds %>% 
  drop_na()  %>% # wirft alle Zeilen raus mit mind. 1 fehlendem Wert
  nrow()
```


```{r}
library(nycflights13)
data(flights)
```


```{r}
flights %>% 
  drop_na() %>% 
  nrow()
```


```{r}
nrow(flights)
```



```{r}
327346/336776
```



```{r}
flights %>% 
  filter(is.na(dep_delay))
```




# set.seed


```{r}
runif(n = 1)
rnorm(n = 1)
```


```{r}
#set.seed(42)
runif(n = 1)
```



# 7.3.4.2


>    What does `na.rm = TRUE` do in mean() and sum()?


```{r}
noten <- c(1, 2, 1, 2, NA)
mean(noten, na.rm = TRUE)
sd(noten, na.rm = TRUE)
sum(noten)
```



# 7.5.1.1

## 1

>   Use what you’ve learned to improve the visualisation of the *departure times* of *cancelled vs. non-cancelled* flights.

```{r}
nycflights13::flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + sched_min / 60
  ) %>% 
  ggplot(mapping = aes(sched_dep_time)) + 
    geom_freqpoly(mapping = aes(colour = cancelled), binwidth = 1/4)
```


```{r}
flights2 <- nycflights13::flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + sched_min / 60
  )

flights_summary <-
  flights2 %>% 
  group_by(cancelled) %>% 
  summarise(sched_dep_time_md = median(sched_dep_time, na.rm = TRUE),
            sched_dep_time_iqr= IQR(sched_dep_time, na.rm = TRUE))

flights2 %>% 
  ggplot() +
  aes(x = cancelled) +
  geom_violin(aes(y = sched_dep_time)) +
  geom_point(data = flights_summary,
             aes(y = sched_dep_time_md),
             color = "red",
             size = 4) +
  geom_errorbar(data = flights_summary,
                aes(ymax = sched_dep_time_md + sched_dep_time_iqr/2,
                    ymin = sched_dep_time_md - sched_dep_time_iqr/2),
                color = "grey40",
                alpha = .7,
                width = .1) +
  labs(caption = "Die roten Punkte stellen den Median dar.")

```


## 2

<https://r4ds.had.co.nz/exploratory-data-analysis.html#exercises-17> 


>   What *variable* in the diamonds dataset is most important for *predicting* the *price* of a diamond? How is that variable "correlated" with *cut*? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?


```{r}
diamonds %>% 
  select(price, carat, depth, table, price, x, y, z) %>% 
  cor()
```


```{r}
p0 <- diamonds %>% 
  ggplot()

p0 +   
  aes(x = cut, y = price) +
  geom_boxplot()
```


```{r}
diamonds %>% 
  ggplot() +
  aes(x = cut, y = carat) +
  geom_boxplot() +
  coord_flip()
```


```{r}
lm1 <- lm(price ~ carat, data = diamonds)

library(modelr)

diamonds2 <- 
  diamonds %>% 
  add_residuals(lm1) %>% 
  add_predictions(lm1)
```


```{r}
p0 +
  aes(x = carat, y = price) +
  geom_point(alpha = .3) +  
  geom_smooth(method = "lm") 

```


```{r}
diamonds2 %>% 
  ggplot() +
  aes(y = resid, x = cut) +
  geom_boxplot() +
  labs(y = "Preis relativ zur Größe des Diamanten")
```



## 6


>   If you have a small dataset, it’s sometimes useful to use *geom_jitter()* to see the relationship between a continuous and categorical variable. The *ggbeeswarm* package provides a number of *methods* similar to geom_jitter(). List them and briefly describe what each one does.


```{r}
library(ggbeeswarm)
```


```{r}
data(mtcars)
```


```{r}
mtcars %>% 
  ggplot() +
  aes(y = mpg, x = hp) +
  geom_jitter(alpha = .5)
```



```{r}
mtcars %>% 
  ggplot() +
  aes(y = mpg, x = hp) +
  geom_beeswarm()
```

```{r}
mtcars %>% 
  ggplot() +
  aes(y = mpg, x = hp) +
  geom_point()
```



```{r}
ggplot(mpg, aes(x = displ, y = cty, color = drv))+
  geom_point(alpha = .5)

ggplot(mpg, aes(x = 1, y = cty, color = drv))+
  geom_beeswarm()

ggplot(mpg, aes(x = 1, y = cty, color = drv))+
  geom_quasirandom()
```











# Tilde, Welle, Kringel

```{r}
 ~  # Alt-N
```


```{r}
y ~ x  # Formelschreibweise
AV ~ UV  # lies: AV ist eine Funktion von UV
```



```{r}
lm(mpg ~ hp, data = mtcars)
```


# Bienenschwarm-Diagramm

```{r}
library(ggbeeswarm)
```


```{r}
ggplot(mpg, aes(x = displ, y = cty, color = drv)) +
  geom_point()
```


```{r}
ggplot(mpg, aes(x = displ, y = cty, color = drv)) +
  geom_point(alpha =.5)
```



```{r}
ggplot(mpg, aes(x = displ, y = cty, color = drv)) +
  geom_jitter()
```


```{r}
ggplot(mpg, aes(x = displ, y = cty, color = drv)) +
  geom_jitter(alpha = .5)
```


```{r}
ggplot(mpg, 
       aes(x = displ, y = cty, color = drv))+
  geom_beeswarm(alpha = .5)
```

```{r}
ggplot(mpg, 
       aes(x = displ, y = cty, color = drv))+
  geom_beeswarm(groupOnX = TRUE)
```



```{r}
ggplot(mpg, 
       aes(x = displ, y = cty, color = drv))+
  geom_beeswarm(groupOnX = FALSE,
                alpha = .5)
```
