---
title: "2021-10-22"
author: "ses"
date: "10/22/2021"
output: html_document
---

# Aufgabe 4

## Aufgabe

2M1. Recall the globe tossing model from the chapter. Compute and plot the grid approximate *posterior distribution* for each of the following sets of observations. In each case, assume a uniform prior for p.

1. WWW 
2. WWWL 
3. LWWLWWW


## Lösung

```{r}
library(tidyverse)
```


```{r}
post <-
  tibble(
    p_grid = seq(0, 1, by = .01),
    prior = 1,
    L1 = dbinom(x = 3, size = 3, prob = p_grid),
    L3 = dbinom(x = 5, size = 7, prob = p_grid),
    post_raw1 = prior * L1,
    post_raw3 = prior * L3,
    post1 = post_raw1 / sum(post_raw1),
    post3 = post_raw3 / sum(post_raw3)    
  )
```


```{r}
post %>% 
  ggplot() +
  aes(x = p_grid, y = post3) +
  geom_point()
```

# Aufgabe 5


## Angabe 

Recall the *globe* tossing model from the chapter. Compute and plot the grid approximate posterior distribution for each of the following sets of observations. In each case, assume a uniform prior for p.

Data:

1. WWW 
2. WWWL 
3. LWWLWWW (5/7)

Now assume a *prior* for p that is equal to *zero* when *p < 0.5* and is a positive constant when p ≥ 0.5. Again compute and plot the grid approximate posterior distribution for each of the sets of observations in the problem just above.


## Lösung

```{r}
post <-
  tibble(
    p_grid = seq(0, 1, by = .01),
    prior = case_when(
      p_grid < 0.5 ~ 0,
      p_grid >= 0.5 ~ 1),
    L1 = dbinom(x = 3, size = 3, prob = p_grid),
    L3 = dbinom(x = 5, size = 7, prob = p_grid),
    post_raw1 = prior * L1,
    post_raw3 = prior * L3,
    post1 = post_raw1 / sum(post_raw1),
    post3 = post_raw3 / sum(post_raw3)    
  )
```




```{r}
post %>% 
  ggplot() +
  aes(x = p_grid, y = post3) +
  geom_point()
```
