---
title: "Modeling your research data: A crash course using R"
author: "ses"
date: "9/14/2021"
output: 
  html_document:
    toc: TRUE
    number_sections: TRUE
---


# Data wrangling

## Exercise


Which three airlines do show the greatest average delay in the first quarter?

(Use the tidyverse approach.)

You'll find the data set (including description) [here](https://vincentarelbundock.github.io/Rdatasets/doc/nycflights13/flights.html).

## Solution


### In pseudo code

take data "flights" THEN
filter month jan to march THEN
select delay THEN
group by airline (carrier) THEN
summarise mean delay
arrange mean delay from biggest to smallest THEN
filter delay the three largest mean delays

### In R code

Some preparation:


```{r message=FALSE}
# make sure to install the packages (one time only!) 
library(nycflights13)  # unless the package is on your machine
data(flights)
library(tidyverse)  # including ggplot, dplyr, ....
```

And now the R code:

```{r}
flights %>% # "and then do command, called the pipe, Contrl-Shift-M"
  filter(month <= 3) %>% 
  select(month, dep_delay, carrier) %>% 
  drop_na() %>%  # drop all rows with missing data, for simplicity
  group_by(carrier) %>% 
  summarise(dep_delay_avg = mean(dep_delay))  %>%   # y = f(x)
  arrange(-dep_delay_avg) %>% 
  mutate(position = 1:16) %>% 
  filter(position <= 3)
```



# Data viz


## Research question 1

Is there a correlation between delay and flight distance?


```{r}
flights %>% 
  select(dep_delay, distance) %>% 
  sample_n(size = 1000) %>%  # for speed
  ggplot() + 
  aes(x = distance, y = dep_delay) +
  geom_point() +
  geom_smooth()
```





## Research question 2

Is there a correlation between delay and flight distance? What is if we split the analysis (correlation) for each airport? (sub group analysis)


```{r}
flights %>% 
  select(dep_delay, distance, origin) %>% 
  sample_n(size = 1000) %>%  # for speed
  ggplot() + 
  aes(x = distance, y = dep_delay, color = origin) +
  geom_point()+
  geom_smooth()
```


## RQ 3

What is the distribution for delay in each month for each airport? Be sure to add a measure of centrality (e.g. median) and for spead (e.g., sd or IQR).


Tipp 1: Boxplots are depicted using the command `geom_boxplot()`!
Tipp 2: Scribble before you code.

*11.40am back here*


```{r}
flights %>% 
  select(dep_delay) %>% 
  drop_na() %>% 
  summarise(mean(dep_delay),
            sd(dep_delay))
```

A rough rule of thumb is:

outlier is greater than mean + 2*sd.

```{r}
uprlimit <- 12.6 + 2*40.2
uprlimit
```

that's ifelse:

ifelse(test, if_yes_do, if_no_do)

```{r}
flights %>% 
  sample_n(1000) %>% 
  mutate(is_outlier = ifelse(dep_delay > 93, 1, 0)) %>% 
  filter(is_outlier == 0) %>% 
  select(dep_delay, month, origin) %>% 
  mutate(month_nominal = factor(month)) %>% 
  ggplot() +
  aes(x = month_nominal, y = dep_delay, color = origin) +
  geom_boxplot()
```

There seems to be a bit less delay in winter times.



# Modelling


## Simple regression


```{r message=FALSE}
library(moderndive)
data(evals)
```



```{r message=FALSE}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
```



## Preparation

```{r}
evals_ch5 <- evals %>%
  select(ID, score, bty_avg, age)
```


## RQ

Is the teaching evaluation *depending* on the beauty of the teacher?


```{r}
evals_ch5 %>% 
  ggplot(aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Relationship between teaching and beauty scores") +  
  geom_smooth(method = "lm", se = FALSE)
```


### Pseudo code

```{r eval = FALSE}
linear_model(output_var ~ input_var,
             data = my_spreadsheet)
```


### Actual code
```{r}
model1 <- lm(score ~ bty_avg,  # y = f(x)
   data = evals_ch5) 

model1
```


```{r}
get_regression_summaries(model1)
```


## Exercise Regression 1: numeric predictor

```{r}
library(moderndive)
data("evals")
```



### Exercise

From [Chapter 5.1.2 ModernDive](https://moderndive.com/5-regression.html#model1table):

**(LC5.2)** Fit a new simple linear regression using `lm(score ~ age, data = evals_ch5)` where age is the new explanatory variable $x$. Get information about the “best-fitting” line from the regression table by applying the `get_regression_table()` function. How do the regression results match up with the results from your earlier exploratory data analysis?



### Solution

```{r}
model2 <- lm(score ~ age, data = evals_ch5)
get_regression_table(model2)
```


```{r}
model2_preds <- get_regression_points(model2)
```




## Exercise Regression 2: Binary predictor


Model `score` (DV/output) as explained by `gender` (IV/input); also visualize the model. Add an explorative data analysis (summary statistics plus visualization). Interpret your results. 
Note that `gender` is a binary (factor) variable.
Using the data `evals` is probably the easiest way to get started (instead of `evals_ch5`).


*approx. 10 Min.*


```{r}
m3 <- lm(score ~ gender, data = evals)
m3
```


```{r}
evals %>% 
  mutate(gender_01 = ifelse(gender == "female", 0, 1)) %>% 
  ggplot() +
  aes(x = gender_01, y = score) +
  geom_point() +
  geom_smooth(method = "lm")
```


```{r}
evals %>% 
  group_by(gender) %>% 
  summarise(mean_per_group = mean(score))
```






## Ex. Regression 3: multinomial predictor

Referring to [ModernDive, 5.2](https://moderndive.com/5-regression.html#model2EDA).


### Ex
```{r}
lm(score ~ rank, data = evals)
```


Model the evaluation score as a function of `rank` of the teacher. Use `evals` as the data. Note that `rank` has 3 nominal values. Then, compute the mean values for each of the 3 groups. 



# Interaction


## Preparing for interaction


###  Ex.

Model `score` by two input variables: `age` and `gender`. But as first step, just *visualize.* Be sure to take the right data frame :-)

Hint: Map `gender` to color :-)

### Solution

Pseudo code:

```{r eval = FALSE}
take the right table THEN
initialize ggplot THEN 
define the 3 aesthetic mappings THEN 
draw geom point THEN 
draw geom smooth (of method "lm")
```


```{r}
ggplot(evals) +
  aes(y = score, x = age, color = gender) +
  geom_point() +
  geom_smooth(method = "lm")
```


This is an *interaction* model.



```{r}
ggplot(evals) +
  aes(y = score, x = age, color = gender) +
  geom_point() +
  geom_parallel_slopes()
```


This is a non-interaction model ("main effects" only).


## Interaction model in R

```{r}
m4 <- lm(score ~ age + gender + age:gender,
         data = evals)
m4
```


